Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-04 robdml.tex
tex

To this end, we propose Graduate Adversary, a family of pseudo-hardness
functions which can be used as the \emph{destination} hardness in HM.
%
Any function that leads to a relatively weak adversary in the early training
phase, and a relatively strong adversary in the late training phase belongs to
this family.
%
As an example, we design a ``Linear Graduate Adversary'' (LGA) pseudo-hardness
function as the linearly scaled negative triplet margin, incorporating a strong
assumption that the \emph{destination} hardness should remain Semihard.
%
%Empirical observation will show the advantage of such pseudo-hardness
%functions.

@@

Besides, a sample triplet will be augmented into a sextuplet (benign and
adversarial examples) during adversarial training, where a \emph{intra-class}
structure can be enforced.
%
Such structure has been neglected by existing methods, but some existing
attacks aims to change the sample rankings in the same class~\cite{advrank}.
%
Hence, we propose a simple \emph{intra-class} structure penalty term for
adversarial training, which is expected to improve robustness.

% evaluation and conclusion
The experiments are conducted on three commonly used DML datasets, namely
CUB-200-2011~\cite{cub200}, Cars-196~\cite{cars196}, and Stanford Online
Product~\cite{sop}.
%
The proposed method overwhelmingly outperforms the state-of-the-art defense in
terms of training cost, robustness, as well as the performance on benign
example retrieval.

\begin{figure}
	\includegraphics[width=\columnwidth]{hmflexible.pdf}
	\vspace{-1.8em}
	\caption{Flexibility of Hardness Manipulation.
	%
	Regular training and min-max adversarial training are two of its boundary cases.
	}
	\label{fig:hmflexible}
\end{figure}

% contributions
In brief, our contributions include:
%
\begin{enumerate}[nosep, noitemsep, leftmargin=*]
	%
	\item {\textit{Hardness Manipulation}} (HM) is proposed as a flexible tool
		to create adversarial triplets with increased hardness for
		subsequent adversarial training.
		%
		It is more efficient than previous adversarial training methods of DML.
		%
	\item \textit{Gradual Adversary} is proposed as a family of pseudo-hardness
		functions for \emph{Hardness Manipulation}, which can balance the
		training objectives during the training process.
		%
		A linear function named Linear Graduate Adversary (LGA) is designed as
		an example.
		%
	\item \textit{Intra-Class Structure} penalty is proposed to further improve
		the robustness of a DML model, while such possibility is neglected by
		existing methods. 
		%
\end{enumerate}

@@

\textbf{Adversarial Attack.}
%
Szegedy \etal~\cite{l-bfgs} find misclassification of DNN can be triggered by
an imperceptible adversarial perturbation to the input image.
%
Ian \etal~\cite{fgsm} speculate the reason is that DNN being locally linear
with respect to the adversarial perturbation.
%
Subsequent studies can compromise the DNNs more effectively, such as
BIM~\cite{i-fgsm}, PGD~\cite{madry}, and APGD~\cite{apgd}, which are
unanimously based on the first-order gradient of the loss, and rely on the
white-box assumption of fully accessible model details.
%
In contrast, query-based black-box attacks~\cite{nes-atk,spsa-atk} are more
practical for real-world scenarios, as they requires the least amount of
information from the model for estimating the gradients.

>>>>

To this end, Graduate Adversary, a family of pseudo-hardness functions is
proposed, which can be used as the \emph{destination} hardness.
%
A function that leads to relatively weak and relatively strong adversarial
examples respectively in the early and late phase of training belongs to this
family.
%
As an example, we design a ``Linear Graduate Adversary'' (LGA)
function as the linearly scaled negative triplet margin, incorporating a strong
prior that the \emph{destination} hardness should remain Semihard
based on our empirical observation.

@@

Additionally, it is noted that a sample triplet will be augmented into a
sextuplet (both benign and adversarial examples) during adversarial training.
%
In this case, the \emph{intra-class} structure can be enforced, has been
neglected by existing methods.
%
Since some existing attacks aim to change the sample rankings in the same
class~\cite{advrank},
%
we propose a simple \emph{intra-class} structure loss term for adversarial
training, which is expected to further improve robustness.

% evaluation and conclusion
Comprehensive experiments are conducted on three commonly used DML datasets,
namely CUB-200-2011~\cite{cub200}, Cars-196~\cite{cars196}, and Stanford Online
Product~\cite{sop}.
%
The proposed method overwhelmingly outperforms the state-of-the-art defense in
terms of training cost, robustness, as well as the performance on benign
examples.

\begin{figure}
	\includegraphics[width=\columnwidth]{hmflexible.pdf}
	\vspace{-1.8em}
	\caption{Flexibility of Hardness Manipulation.
	%
	Regular training and min-max adversarial training are its boundary cases.
	}
	\label{fig:hmflexible}
\end{figure}

% contributions
In brief, our contributions include:
%
\begin{enumerate}[nosep, noitemsep, leftmargin=*]
	%
	\item {\textit{Hardness Manipulation}} (HM) is proposed as a flexible and
		efficient tool to create adversarial example triplets for subsequent
		adversarial training of DML model.
		%
	\item \textit{Linear Gradual Adversary} (LGA) is proposed as a
		pseudo-hardness function for HM, which incorporates our empirical
		observations and can balance the training objectives during the
		training process.
		%
	\item \textit{Intra-Class Structure} (ICS) loss term is proposed to further
		improve the DML model robustness, while such possibility is neglected
		by existing methods. 
		%
\end{enumerate}

@@

\textbf{Adversarial Attack.}
%
Szegedy \etal~\cite{l-bfgs} find misclassification of DNN can be triggered by
an imperceptible adversarial perturbation to the input image.
%
Ian \etal~\cite{fgsm} attribute the reason to DNN being locally linear with
respect to the adversarial perturbation.
%
Subsequent first-order gradient-based methods can compromise the DNNs more
effectively under the white-box assumption~\cite{i-fgsm,madry,apgd}.
%
In contrast, black-box attacks have been explored by query-based
methods~\cite{nes-atk,spsa-atk} and transferrability-based methods~[], which
are more practical for real-world scenarios.
